### 2023-12-13 experiments/notebooks/poc.qmd
rerunning POC chart

@@

@@

```{julia}
method = :simple_inductive
cov = .95
conf_model = conformal_model(clf; method=method, coverage=cov)
mach = machine(conf_model, table(permutedims(X)), labels)
fit!(mach)
Serialization.serialize(joinpath(output_path,"poc_model.jls"), mach)
```

```{julia}
#| echo: false
niter = 1000
jem = mach.model.model.jem
_batch_size = mach.model.model.batch_size
X = Float32.(matrix(X))
if typeof(jem.sampler) <: ConditionalSampler

@@

@@

temp = 0.3
p0 = Plots.contourf(mach.model, mach.fitresult, permutedims(X), labels; plot_set_size=true, zoom=0, temp=temp)
p1 = Plots.contourf(mach.model, mach.fitresult, permutedims(X), labels; plot_set_loss=true, zoom=0, temp=temp)
p2 = Plots.contourf(mach.model, mach.fitresult, permutedims(X), labels; plot_classification_loss=true, zoom=0, temp=temp, clim=nothing, loss_matrix=ones(2,2))
plt = display(Plots.plot(p0, p1, p2, size=(1400,320), layout=(1,3)))
savefig(joinpath(output_images_path, "poc_set_size.png"))
```

```{julia}
Random.seed!(1234)

λ₁ = 0.1
λ₂ = 0.4
λ₃ = 0.4
Λ = [λ₁, λ₂, λ₃]

@@

ces = Dict{Any,Any}()
plts = []
for (name, generator) in generator_dict
    ce = generate_counterfactual(
        x_factual, target, counterfactual_data, M, generator;
        initialization=:identity, 
        converge_when=:generator_conditions,
    )
    plt = Plots.plot(
        ce, title=name, alpha=0.2, cbar=false, 
        axis=nothing, length_out=10, contour_alpha=1.0,
    )
    if name == "ECCCo"
        _X = distance_from_energy(ce, return_conditionals=true)
        Plots.scatter!(
            _X[1,:],_X[2,:], color=:purple, shape=:star5, 
            ms=10, label="x̂|$target", alpha=0.1
        )
    end
    push!(plts, plt)
    ces[name] = ce
end
plt = Plots.plot(plts..., size=(500,520))
display(plt)
savefig(plt, joinpath(output_images_path, "poc.png"))
```

```{julia}
using Colors
col_pal = palette(:seaborn_colorblind)
Random.seed!(1234)
using CounterfactualExplanations.Generators: ∇

λ₁ = 0.1
λ₂ = 0.5
λ₃ = 0.5
Λ = [λ₁, λ₂, λ₃]
η = 0.01

@@

generator_dict = OrderedDict(
    "Wachter" => WachterGenerator(λ = 0.3, opt=opt),
    "ECCCo (no EBM)" => ECCCoGenerator(λ = [λ₁,λ₂,0.0], opt=opt),
    "ECCCo (no CP)" => ECCCoGenerator(λ = [λ₁,0.0,λ₃], opt=opt),
    "ECCCo" => ECCCoGenerator(λ = Λ, opt=opt),
)

@@

    # Gradient field:
    u = []
    v = []
    for (x, y) in inputs
        g = -loss_grads(generator, M, ce, [x, y][:,:])
        push!(u, η * g[1])
        push!(v, η * g[2])
    end
    Plots.quiver!(x1, x2, quiver=(u, v), color=col_pal[5])
    # arrow0!.(x1, x2, u, v; as=0.2, lw=1.0,lc=col_pal[5], la=1)
    push!(plts, plt)
    ces[name] = ce
end
plt = Plots.plot(plts...; size=(panel_height*length(plts),panel_height), layout=(1,length(plts)), dpi=300)
# plt = Plots.plot(plts..., size=(1000,250), layout=(1,4), dpi=300)
display(plt)
savefig(plt, joinpath(output_images_path, "poc_gradient_fields.png"))
```

```{julia}
ce = ces["ECCCo"]
using CounterfactualExplanations.Generators: ℓ

# loss
function f1(x)
    _ce = deepcopy(ce)
    _ce.s′ = x
    return - ℓ(_ce.generator, _ce)
end

# cost
function f2(x)
    _ce = deepcopy(ce)
    _ce.s′ = x
    λ = _ce.generator.λ[1]
    _loss = _ce.generator.penalty[1]
    return - λ * _loss(_ce)
end

# set size
function f3(x)
    _ce = deepcopy(ce)
    _ce.s′ = x
    λ = _ce.generator.λ[2]
    _loss = _ce.generator.penalty[2]
    return - λ * _loss(_ce)
end

# distance from energy
function f4(x)
    _ce = deepcopy(ce)
    _ce.s′ = x
    λ = _ce.generator.λ[3]
    _loss = _ce.generator.penalty[3]
    return - λ * _loss(_ce)
end

# Helper function:
meshgrid(x, y) = (repeat(x, outer=length(y)), repeat(y, inner=length(x)))

xlims, ylims = extrema(X, dims=2)
xrange = range(xlims..., length=10)
yrange = range(ylims..., length=10)

x1, x2 = meshgrid(xrange, yrange)
inputs = zip(x1, x2)

u = []
v = []

for (x, y) in inputs
    push!(u, scale * gradient(f1, [x, y][:,:])[1][1])
    push!(v, scale * gradient(f1, [x, y][:,:])[1][2])
end

Plots.plot(xlims=xlims, ylims=ylims)
Plots.scatter!(counterfactual_data)
Plots.quiver!(x1, x2, quiver=(u, v))
```

>>>>

## Data

@@

## Model

@@

```{julia}
method = :simple_inductive
cov = .95
conf_model = conformal_model(clf; method=method, coverage=cov)
mach = machine(conf_model, table(permutedims(X)), labels)
fit!(mach)
```

## Conditional Samples

```{julia}
#| echo: false

using MLJBase
niter = 1000
jem = mach.model.model.jem
_batch_size = mach.model.model.batch_size
X = Float32.(MLJBase.matrix(X))
if typeof(jem.sampler) <: ConditionalSampler

@@

## Losses and Penalties

@@

temp = 0.3
p0 = Plots.contourf(mach.model, mach.fitresult, permutedims(X), labels; plot_set_size=true, zoom=0, temp=temp)
p1 = Plots.contourf(mach.model, mach.fitresult, permutedims(X), labels; plot_set_loss=true, zoom=0, temp=temp)
plt = display(Plots.plot(p0, p1, size=(800,320), layout=(1,2)))
```

## Counterfactuals

Compared to other generators:

```{julia}
using OrderedCollections
Random.seed!(1234)

λ₁ = 0.1
λ₂ = 0.4
λ₃ = 5.0
Λ = [λ₁, λ₂, λ₃]

@@

ces = Dict{Any,Any}()
plts = []
for (name, generator) in generator_dict
    ce = generate_counterfactual(
        x_factual, target, counterfactual_data, M, generator;
        initialization=:identity, 
        converge_when=:generator_conditions,
    )
    plt = Plots.plot(
        ce, title=name, alpha=0.2, cbar=false, 
        axis=nothing, length_out=10, contour_alpha=1.0,
    )
    if name == "ECCCo"
        _X = distance_from_energy(ce, return_conditionals=true)
        Plots.scatter!(
            _X[1,:],_X[2,:], color=:purple, shape=:star5, 
            ms=10, label="x̂|$target", alpha=0.1
        )
    end
    push!(plts, plt)
    ces[name] = ce
end
plt = Plots.plot(plts..., size=(500,520))
display(plt)
```

```{julia}
using Colors
panel_height = 200
col_pal = palette(:seaborn_colorblind)
Random.seed!(1234)
using CounterfactualExplanations.Generators: ∇

λ₁ = 0.2
λ₂ = 0.5
λ₃ = 3.0
Λ = [λ₁, λ₂, λ₃]
η = 0.01

@@

generator_dict = OrderedDict(
    "Wachter" => WachterGenerator(λ = 0.7, opt=opt),
    "ECCCo (no EBM)" => ECCCoGenerator(λ = [λ₁,λ₂,0.0], opt=opt),
    "ECCCo (no CP)" => ECCCoGenerator(λ = [λ₁,0.0,λ₃], opt=opt),
    "ECCCo" => ECCCoGenerator(λ = Λ, opt=opt),
)

@@

    # Gradient field:
    u = []
    v = []
    for (x, y) in inputs
        g = -loss_grads(generator, M, ce, [x, y][:,:])
        push!(u, η * g[1])
        push!(v, η * g[2])
    end
    Plots.quiver!(x1, x2, quiver=(u, v), color=col_pal[5])
    # arrow0!.(x1, x2, u, v; as=0.2, lw=1.0,lc=col_pal[5], la=1)
    push!(plts, plt)
    ces[name] = ce
end
plt = Plots.plot(plts...; size=(panel_height*length(plts),panel_height), layout=(1,length(plts)), dpi=300)
# plt = Plots.plot(plts..., size=(1000,250), layout=(1,4), dpi=300)
display(plt)
```
